Complete Algebraic File Synchronization

### 2015-12-04 2015/paper.tex
Split source into sections

\newcommand{\setv}{\mathcal{V}} % Set of values
\newcommand{\setvx}[1]{\mathcal{V}_{#1}} % Set of values with custom index
\newcommand{\setf}{\setvx{F}} % Set of file values
\newcommand{\setd}{\setvx{D}} % Set of directory values
\newcommand{\setb}{\setvx{\empt}} % Set of empty values
\newcommand{\setn}{\mathcal{N}} % Set of nodes / paths
\newcommand{\empt}{\bot}
\newcommand{\parent}{\mathtt{parent}}
\newcommand{\topnode}{\mathtt{None}} % return value of parent()
\newcommand{\fsbroken}{\mathtt{Broken}} % broken FS value
\newcommand{\FS}{\Phi} % a filesystem
\newcommand{\GS}{\Xi} % another filesystem
\newcommand{\nn}{n^*} % parent node

\newcommand{\cbrk}{\mathtt{break}} % 'break' command
\newcommand{\ccharb}{\empt}
\newcommand{\ccharf}{\mathtt{F}}
\newcommand{\cchard}{\mathtt{D}}
\newcommand{\vald}{d} % Directory value
\newcommand{\valdi}{d_1}
\newcommand{\valdii}{d_2}
\newcommand{\valf}{f} % File value
\newcommand{\valvx}{v_X} % Value of type X
\newcommand{\valvy}{v_Y} % Value of type Y
\newcommand{\valvw}{v_W}
\newcommand{\valvr}{v_R}
\newcommand{\typeset}{\{\cchard,\ccharf,\ccharb\}} % Set of types in commands
\newcommand{\caa}[2]{\langle{#1,#2}\rangle}
\newcommand{\caaa}[3]{\langle{#1,#2,#3}\rangle}
\newcommand{\caaaa}[4]{\langle{#1,#2,#3,#4}\rangle}
\newcommand{\cbb}{\caa{\ccharb}{\ccharb}}
\newcommand{\cbba}[1]{\caaa{\ccharb}{\ccharb}{#1}}
\newcommand{\cbf}{\caa{\ccharb}{\ccharf}}
\newcommand{\cbfaa}[2]{\caaaa{\ccharb}{\ccharf}{#1}{#2}}
\newcommand{\cbd}{\caa{\ccharb}{\cchard}}
\newcommand{\cbdaa}[2]{\caaaa{\ccharb}{\cchard}{#1}{#2}}
\newcommand{\cfb}{\caa{\ccharf}{\ccharb}}
\newcommand{\cfba}[1]{\caaa{\ccharf}{\ccharb}{#1}}
\newcommand{\cff}{\caa{\ccharf}{\ccharf}}
\newcommand{\cfd}{\caa{\ccharf}{\cchard}}
\newcommand{\cfdaa}[2]{\caaaa{\ccharf}{\cchard}{#1}{#2}}
\newcommand{\cdb}{\caa{\cchard}{\ccharb}}
\newcommand{\cdba}[1]{\caaa{\cchard}{\ccharb}{#1}}
\newcommand{\cdf}{\caa{\cchard}{\ccharf}}
\newcommand{\cdfaa}[2]{\caaaa{\cchard}{\ccharf}{#1}{#2}}
\newcommand{\cdd}{\caa{\cchard}{\cchard}}
\newcommand{\cdda}[1]{\caaa{\cchard}{\cchard}{#1}}
\newcommand{\cxy}{\caa{X}{Y}}
\newcommand{\cxyaa}[2]{\caaaa{X}{Y}{#1}{#2}}
\newcommand{\cxynv}{\caaaa{X}{Y}{n}{\valvy}}
\newcommand{\cxynnv}{\caaaa{X}{Y}{\nn}{\valvy}}
\newcommand{\cxw}{\caa{X}{W}}
\newcommand{\cxwnv}{\caaaa{X}{W}{n}{\valvw}}
\newcommand{\czw}{\caa{Z}{W}}
\newcommand{\czwnv}{\caaaa{Z}{W}{n}{\valvw}}
\newcommand{\czwnnv}{\caaaa{Z}{W}{\nn}{\valvw}}
\newcommand{\czwmv}{\caaaa{Z}{W}{m}{\valvw}}
\newcommand{\cqr}{\caa{Q}{R}}
\newcommand{\cqrmv}{\caaaa{Q}{R}{m}{\valvr}}
\newcommand{\cqrov}{\caaaa{Q}{R}{o}{\valvr}}

\newcommand{\cc}{\circ} % Command / sequence concatenation
\newcommand{\descendant}{\prec}
\newcommand{\descendantEq}{\preceq}
\newcommand{\ancestor}{\succ}
\newcommand{\ancestorEq}{\succeq}

\newcommand{\eqext}{\sqsubseteq} % equivalent to
\newcommand{\eqnrw}{\sqsupseteq} % extended by
\newcommand{\nequiv}{\not\equiv}
\newcommand{\indep}{\mathrel{\wr\wr}} % Independent commands, sequences
\newcommand{\unrel}{\indep} % Unrelated nodes
\newcommand{\nindep}{\,{\centernot{\wr\wr}}\,} % not \indep
\newcommand{\nunrel}{\nindep} % not \unrel

\newcommand{\workssign}{\mathbf{w}}
\newcommand{\works}[1]{\workssign({#1})}
\newcommand{\worksc}[2]{\workssign({#1}\mathrel{|}{#2})}

\newcommand{\typeeq}{\stackrel{t}{=}} % Type equality
\newcommand{\ntypeeq}{\centernot{\stackrel{t}{=}}}

\newcommand{\emptyseq}{\lambda} % empty sequence

\newcommand{\orderset}[1]{\vec{\wp}({#1})}
\newcommand{\seqset}[1]{\underline{#1}} % Set of sequences

\newcommand{\recchar}[3]{{#1}^{#3}_{\mathcal{R}|{#2}}}
\newcommand{\reca}{\recchar{A}{B}{}} % Reconciled from A
\newcommand{\recb}{\recchar{B}{A}{}} % Reconciled from B

\newcommand{\acb}{(A\cap B)}
\newcommand{\acbi}{\acb^{-1}}
\newcommand{\amb}{(A\setminus B)}
\newcommand{\bma}{(B\setminus A)}

\theoremstyle{definition}

\newtheorem{mydef}{Definition}
\crefname{mydef}{definition}{definitions}

\newtheorem{myax}{Rule}
\crefname{myax}{rule}{rules}

\newtheorem{mylem}{Lemma}
\crefname{mylem}{lemma}{lemmas}

\newtheorem{myth}{Theorem}
\crefname{myth}{theorem}{theorems}

\newtheorem{mycor}{Corollary}
\crefname{mycor}{corollary}{corollaries}

@@

\section{Definition of a Filesystem}

We model a filesystem using a function $\FS$ with a set of nodes (filesystem paths) $\setn$ as its domain,
and a set of possible contents $\setv$ as its codomain:
\[ \FS : \setn \rightarrow \setv \] 
In our model, $\setn$ contains all possible nodes, including the ones where there is no file or directory,
and $\setv$ contains a special element, $\empt$, which is the value of $\FS$ at an empty node.
We consider the contents of files, as well as any meta-information of files
and directories (e.g. permissions or flags) part of the values in $\setv$.

The nodes form a disjoint union of rooted directed trees.
\begin{mydef}
The function $\parent(n)$ returns the parent node of $n$, or
returns $\topnode$ if $n$ is the root of a tree.
\end{mydef}

Every filesystem has a so-called \textbf{tree property}, which means that
if the filesystem is not empty at a node, and the node has a parent,
then there must be a directory at the parent node.

To model this, in $\setv$ we distinguish between files ($\setf$) and directories ($\setd$), that is,
if $\setb = \{\empt\}$ then:
\[ \setv = \setb \cup \setf \cup \setd \]
The tree propety can then be defined as
\begin{align*}
\forall n\in\setn: &\FS(n) \neq \empt \\ % TODO alignment problem
&\quad\wedge \parent(n) \neq \topnode \\
&\Rightarrow \FS(\parent(n)) \in \setd 
\end{align*}
This means that as we move down from the root of a tree of nodes,
the types of values we encounter in the filesystem can only change according to the following
transition diagram:

\input{type-transition}

In this paper, 
we write $n$ and $m$ for nodes in $\setn$,
and use lowercase Greek letters to denote commands. % TODO but very rarely
We use $\valf$ for an arbitrary element of $\setf$, and $\vald$ for an arbitrary element
of $\setd$. $TODO$ is a value from $\setv$ .

\section{Introducing commands}

Next we need to define commands on the filesystem.

@@

...

Let us consider what kind of information may be encoded in the commands.
A usual set of commands, based on the most frequent tools implemented by filesystems,
might be $create(n,V)$, $edit(n,V)$ and $remove(n)$ where $n\in\setn$ and $V\in\setv$ (but $V\neq\empt$).
Regarding their output, that is, the state of the filesystem at $n$
after applying the command,
we know that after $create$ or $edit$, $\FS(n)\neq\empt$, whereas after $remove$,
$\FS(n)$ will be $\empt$. 
However, from \cite{NREC:alg} we know that a useful set of axioms
will in some cases need to distinguish between edits that result in directories ($edit(n,D)$) and
ones that result in files ($edit(n,F)$), and treat them as separate commands.
Indeed, Bill Zissimopoulos suggested (\cite{BZ}) that extending this distinction to more commands
utlimately simplifies
the definition of conflicting commands, as our model will then able to predict the behavior of commands
more precisely.
In other words, encoding the type of the output ($D$, $F$ or $\empt$) in the commands is definitely useful.
At the same time, there is never any need to encode information about the
\emph{exact} output value of a command,
as the success or failure of filesystem commands only depend on the types of values in the filesystem.

Notice, however, that the commands listed above also encode some information about 
their input, the filesystem
\emph{before} the command is applied; namely, $create(n,V)$ requires that there are no files
or directories at $n$, while $edit(n,V)$ and $remove(n)$ require the opposite.
This creates an arbitrary asymmetry where
there is now more information encoded into commands about their output than about their input.
In line with the aim to encode as much information as possible in the commands, % TODO Where is this mentioned?
and in order to resolve this asymmetry, we propose a set of commands that encode
the type of the original state of $\FS(n)$ as well.
(Some real-life filesystem commands like $rmdir$ do this already.)

% TODO: tuple / quadruple
We therefore propose to have a command for each pair of types.
For want of a better system, we will name our commands by concatenating
two of $\cchard$, $\ccharf$ and $\ccharb$, 
where the left sign denotes the type of value
required in the filesystem before the command is applied, and the right sign notes the type
afterwards. For example, $mkdir(n,D)$ is $\cbdaa{n}{D}$, and $rmdir(n)$ is $\cdba{n}$.

A command can only succeed if the original value at node $n$ has a type that matches
the input type of the command. If this is not the case, or if the resulting
filesystem no longer has the tree property, then we say that the command
{\em breaks} the filesystem. Broken filesystems are considered equal
(but not equal to any working filesystem), and we note them by $\FS=\fsbroken$.

So that we could reason about sequences of commands that break every filesystem
in the algebra, we introduce the command $\cbrk$ that simply breaks every filesystem.

We note that a command or a sequence of commands is applied to a filsystem
by prefixing the command or sequence to it, for example: $\cbrk\FS$, $\cbdaa{n}{\vald}\FS$, 
or $S\FS$ if $S$ is a sequence of commands.

\begin{mydef}
Using $\FS[V/n]$ we denote a filesystem derived from $\FS$:
\[ \FS[V/n](m) :=
   \begin{cases}
   V &\mbox{if~} m=n\\
   \FS(m) &\mbox{otherwise.}
   \end{cases}
\]
\end{mydef}

The exact behavior of our commands is as follows:
\begin{itemize}
\item $\cbrk\FS = \fsbroken$
\item
All other commands have the generic form $\cxynv$ where
$X$ and $Y$ are types in $\typeset$,
$n\in\setn$ and $\valvy$ is a value of the appropriate type from $\setvx{Y}$.
At times we omit $\valvy$ if there is only one suitable value.
\[ \cxynv\FS = 
   \begin{cases}
   \fsbroken &\mbox{if~} \FS=\fsbroken\\
   \fsbroken &\mbox{if~} \FS(n)\not\in\setvx{X}\\
   \fsbroken &\mbox{if~} \FS[V/n] \mbox{~violates the tree property}\\
   \FS[V/n] &\mbox{otherwise.}
   \end{cases}
\]
\end{itemize}

In general, when we describe multiple or unknown commands, we may substitute one or both
of their types with variables, as we did with $\cxynv$. In this context,
if we write that $\cxy=\czw$, we mean that the input and output types % TODO needed?
of the commands are the same ($X=Z$ and $Y=W$), while a full equiality 
($\cxynv=\czwmv$)
implies
that their output value is also the same.

% Simplifications
% ---------------

For reasons also listed in \cite{NREC:alg}, in this paper we will not consider
a $move$ or $rename$ command. Regarding the theoretical reasoning we aim to follow,
this turns out to be useful because these would be the only commands that affect
filesystems at two nodes at once, therefore describing 
the dependencies affecting these commands
would call for a more complicated model.
From a pragmatic perspective, this restriction does not mean that in an application
implementing conflict resolution using the algorithm described here would not be
able to handle renames by pre- and post-processing changes in the filesystem to
discover them, which (especially when a rename is combined with changes to the content)
is a non-trivial problem in itself.

Another simplification we will assume in this paper is regarding directories.
As Bill Zissimopoulos pointed out (\cite{BZ}), we often do not want to consider metadata stored in
directories during synchronization. In other words, the contents of directories are all equal,
which can be modelled by assuming that $|\setd|=1$.
% This also creates a convenient symmetry where there is only a single value
% for each type that can be repeated as we move from a path to its child ($D$ and $\empt$).

\section{Describing the behavior of commands}

% Definitions for command pairs and ordering
% ------------------------------------------

So that we could describe the effects of commands on a filesystem, let us introduce some notation
and note some observations.

The commands we aim to investigate do not occur in isolation:
for example, when comparing filesystems to determine the commands needed to transform one into the other,
we receive a set of commands with uncertain ordering.
Applying commands to a filesystem, however, happens in time,
so for this we need a sequence of commands.
We use $\cc$ to concatenate commands to form a sequence, or concatenate sequences to form a longer sequence:
$S=\cbdaa{n}{\vald}\cc\cbfaa{m}{\valf}$, or $S'=S\cc\cbba{o}$.
We write $\emptyseq$ for the empty sequence.

TODO: Sequences of commands, in the usual way,  form a free
semigroup

The following concepts allow us to investigate sequences independently of filesystems.
In the definitions, $A,B,S,T$ stand for arbitrary sequences of commands.
These relations echo the ones defined in \cite{NREC:alg}.

\begin{mydef}
% Relation in the algebra
$A\equiv B$ means that $A$ and $B$ are equivalent,
% intended interpretation:
that is, they behave in the same way on
all filesystems: $\forall \FS: A\FS=B\FS$.
It is a reflexive, commutative and transitive relation,
and clearly
% inference rule:
$ A\equiv B \Rightarrow S\cc A\cc T\equiv S\cc B\cc T $.
\end{mydef}

\begin{mydef}
% Relation in the algebra
$A\eqext B$ means that $B$ extends $A$,
% intended interpretation:
that is, they behave in the same way
on any filesystem $A$ does not break:
$\forall \FS: A\FS\neq\fsbroken\Rightarrow A\FS=B\FS$.
It is a reflexive and transitive relation,
and we can see that
% inference rules:
$ A\eqext B \Rightarrow S\cc A\cc T\eqext S\cc B\cc T$
and
$ A\equiv B \Rightarrow A\eqext B$.
\end{mydef}

Let us also define the following relations between nodes:

\begin{mydef}{$n\descendant m$.}
We write $n\descendant m$ if $n$ is the ancestor of $m$ ($n=\parent^n(m)$). % TODO (the transitive closure of the $\parent$ relation)
We write $n\descendantEq m$ if $n\descendant m$ or $n=m$.
\end{mydef}

\begin{mydef}{$n\unrel m$.}
We write $n\unrel m$, or ``$n$ and $m$ are unrelated'' if $n\not\descendantEq m$ and $n\not\ancestorEq m$.
\end{mydef}

% Command pairs - Axioms
% ----------------------

\section{Investigating command pairs}\label{section_axioms}

Our aim is to derive information about the effects of sequences
of commands independently of that actual filesystems.
In order to make this possible, we investigate the smallest building
blocks of sequences: pairs of commands that act on a filesystem directly one after the other.
This approach is useful as there are a limited number of command pairs,
because, as we argued above, we can disregard the exact output values of commands apart from their type,
and we can also abstract the relationship between the nodes in the two commands
to a finite number of cases.

We methodically investigated of all possibilities using a computer program
to determine
which pairs of commands cause errors all the time,
which can be simplified to one or no commands, and which can be reversed
without any change in their overall effect.
% TODO: Footnote: URL
Our basis for this investigation was the model of filesystems introduced in this paper.
These properties of command pairs are crucial as they determine
how a set of commands can be re-ordered to be applied to a filesystem
during synchronization, and what commands will never be compatible.

Below we list a number of statements derived using this method.
Although they can be proven by investigating all relevant possibilities
in a filesystem, we refer to these as axioms because in a secondary algebraic model
that describes commands without reference to an underlying filesystem,
these would function as meta-rules generating statements we would accept as true.
See section (TODO) for more about such a model.

Pairs of commands in general have the form
\[ \cxynv\cc  \czwmv \]
where $X,Y,Z,W\in\typeset$, $n,m\in\setn$, 
and values are of the appropriate type: 
$\valvy\in\setvx{Y}$ and $\valvw\in\setvx{W}$.

\begin{myax}\label{ax_separate_commute}
Commands on unrelated nodes commute:
$\cxynv\cc\czwmv \equiv \czwmv\cc\cxynv$ where $n\unrel m$.
\end{myax}

\begin{myax}\label{ax_separate_nobreaks}
Commands on unrelated nodes do not break every filesystem:
$\cxynv\cc\czwmv \nequiv \cbrk$ where $n\unrel m$.
\end{myax}

\begin{myax}\label{ax_same_breaks}
Commands on the same node break every filesystem if their types are incompatible:
$\cxynv\cc\czwnv \equiv \cbrk$ where $Y\ne Z$.
\end{myax}

\begin{myax}\label{ax_same_emptyseq}
Commands on the same node simplify:
$\cxynv\cc\czwnv \eqext \emptyseq$ where $Y=Z$ and $X=W=\ccharb$ 
or $X=W=\cchard$.
\end{myax}

\begin{myax}\label{ax_same_singlec}
Commands on the same node simplify:
$\cxynv\cc \czwnv \equiv \cxwnv$ where $Y=Z$ and $\cxw\neq\cbb$ and $\cxw\neq\cdd$.
\end{myax}

\begin{myax}\label{ax_distantrel_breaks}
Commands on distant relatives break all filesystems:
$\cxynnv\cc\czwnv \equiv \cbrk$
and $\czwnv\cc\cxynnv \equiv\cbrk$
where $\nn\descendant n$ and $\nn\neq\parent(n)$ and $\cxy\neq\cdd$ and $\czw\neq\cbb$.
\end{myax}

\begin{mydef}{Construction pair.}
A pair of commands on nodes $\nn$ and $n$ is a construction pair if $\nn=\parent(n)$ 
and the commands are one
of the following:
   \begin{gather*}
   % bd>bf
            \cbdaa{\nn}{\vald}\cc  \cbfaa{n}{\valf} \\
   % bd>bd
            \cbdaa{\nn}{\valdi}\cc  \cbdaa{n}{\valdii} \\
   % fd>bf
            \cfdaa{\nn}{\vald}\cc  \cbfaa{n}{\valf} \\
   % fd>bd
            \cfdaa{\nn}{\valdi}\cc  \cbdaa{n}{\valdii}
   \end{gather*}
\end{mydef}

\begin{myax}\label{ax_directchild_breaks}
All other commands on a child break every filesystem:
$\cxynnv\cc\czwnv \equiv \cbrk$ where $\nn=\parent(n)$ and $\cxy\neq\cdd$ and $\czw\neq\cbb$
and the pair is not a construction pair.
\end{myax}

\begin{mydef}{Destruction pair.}
A pair of commands on nodes $n$ and $\nn$ is a destruction pair if $\parent(n)=\nn$ and the commands are one
of the following:
   \begin{gather*}
   % fb<db ??
            \cfba{n}\cc  \cdba{\nn} \\
   % fb<df ??
            \cfba{n}\cc  \cdfaa{\nn}{\valf} \\
   % db<db ??
            \cdba{n}\cc  \cdba{\nn} \\
   % db<df ??
            \cdba{n}\cc  \cdfaa{\nn}{\valf}
   \end{gather*}
\end{mydef}

\begin{myax}\label{ax_directparent_breaks}
All other commands on a parent break every filsystem:
$\cxynv\cc\czwnnv \equiv \cbrk$ where $\parent(n)=\nn$ and $\cxy\neq\cbb$ and $\czw\neq\cdd$
and the pair is not a destruction pair.
\end{myax}

\begin{mydef}{Assertion command.}
A command is an assertion command if
for every filesystem it either breaks it or leaves it in the same state.
In other words, assertion commands have the same input and output type,
and there must also be only one possible value for their output.
Accordingly, in our current algebra, $\cbba{n}$ and $\cdda{n}$ are the only two assertion commands,
and we do not note their output value as it is determined.
\end{mydef}

\begin{myax}\label{ax_child_assert}
An assertion command can be added on a descendant node:
$\cbba{n}\cc\cxynnv \equiv \cxynnv \equiv \cxynnv\cc\cbba{n}$ 
where $\nn\descendant n$ and $\cxy\neq\cdd$.
\end{myax}

\begin{myax}\label{ax_parent_assert}
An assertion command can be added on an ancestor node:
$\cdda{\nn}\cc\cxynv \equiv \cxynv \equiv \cxynv\cc\cdda{\nn}$ 
where $\nn\descendant n$ and $\cxy\neq\cbb$.
\end{myax}

\begin{myax}\label{ax_assert}
Assertion commands can be removed:
$\cxynv \eqext \emptyseq$ where $\cxy=\cbb$ or $\cxy=\cdd$.
\end{myax}

@@

% Works
% -----

\begin{mydef}
% Unsure if this relation should be in the algebra.
For a set of $k$ sequences
$\works{A_1,A_2,\ldots,A_k}$ means that 
$A_1,A_2,\ldots,A_k$ work at the same time, that is,
% intended interpretation:
\[\exists \FS: A_1\FS\neq\fsbroken \wedge \cdots \wedge A_k\FS\neq\fsbroken\]
As the sequences form a set, their order is irrelevant.
% inference rules:
We also know that $\works{A} \Leftrightarrow A\nequiv \cbrk$. 
\end{mydef}

\begin{mydef}
For two sets of sequences, $\worksc{A_1,A_2,\ldots,A_k}{B_1,B_2,\ldots,B_l}$ means that 
all of $A_1,\ldots,A_k$ work where all of $B_1,\ldots,B_l$ work,
that is,
% intended interpretation:
\begin{align*}
\forall \FS:&\\ 
&B_1\FS\neq\fsbroken \wedge \cdots \wedge B_k\FS\neq\fsbroken\\
&\Rightarrow\\
&A_1\FS\neq\fsbroken \wedge \cdots \wedge A_l\FS\neq\fsbroken
\end{align*}
As above, the order of the sequences in the sets is not relevant.
\end{mydef}

\begin{mydef}{Sets of sequences.}
For brevity, we will use $\seqset{S}$ to note an arbitrary set of sequences,
and $A\cc\seqset{S}$ to note the set of sequences $\{A\cc S|S\in\seqset{S}\}$.
\end{mydef}

It is easy to see that the following corollaries are true:

% An inference rule in the algebra
% $\worksc{X}{X}$ is always true for any set of sequences $X$,

\begin{mycor}
% An inference rule in the algebra
$B\eqext A \Rightarrow \worksc{A}{B}$.
\end{mycor}

\begin{mycor}\label{worksextpostfix}
% An inference rule in the algebra
$\forall A,S: \worksc{A}{A\cc S}$, that is, if a sequence works, its prefix also works.
\end{mycor}

\begin{mycor}\label{workschained}
% An inference rule in the algebra
$\workssign$ can be chained:
if $\seqset{A'}\subset\seqset{A}$, then
$ \worksc{\seqset{C}}{\seqset{A'}} \wedge \worksc{\seqset{A}}{\seqset{B}} \Rightarrow \worksc{\seqset{A}\cup\seqset{C}}{\seqset{B}}$.
\end{mycor}

We also prove the following lemmas.

\begin{mylem}\label{unrelated_is_independent}
Two different commands that are not assertion commands are independent iff the nodes they change are unrelated:
\[ \cxynv\indep\czwmv \Leftrightarrow n\unrel m \]
if $\cxynv\neq\czwmv$ and none of $\cxy$ and $\cxw$ is $\cbb$ or $\cdd$.
\end{mylem}
\begin{proof}
This is easy to show based on the output of the software that investigates the behavior of command pairs,
and which we used to derive the \namecrefs{ax_separate_commute} in \cref{section_axioms}.
However, this proposition can also be derived from the \namecrefs{ax_separate_commute} already listed
in the following way.

We proceed in two steps. First, we note that
from \cref{ax_separate_commute,ax_separate_nobreaks} we know that
if $n\unrel m$, then $\cxynv$ and $\czwmv$ must necessarily be independent, and therefore
$ \cxynv\nindep\czwmv \wedge n\unrel m $ cannot be the case.

Next, we prove that
\[ \cxynv\indep\czwmv \wedge n\nunrel m \] is also impossible as it would lead to contradiction.
As $n\nunrel m$
we know that $n=m$, $n\descendant m$ or $m\descendant n$.

If $n=m$, then from \cref{ax_same_breaks} we know $Y=Z$ and $W=X$
as otherwise the commands, in one order or the other, would break all filesystems
and they could not be independent.
The commands are not assertion commands, and so from \cref{ax_same_singlec}
we know that $\cxynv\cc\czwnv\equiv\czwnv\cc\cxynv$ can only happen if $Y=W$ and $\valvy=\valvw$.
Therefore $X=Y=Z=W$ which is only possible if both commands are $\cff$ and have the same
output value. This, however, contradicts our assumption that they are different.

If $n\descendant m$ or $m\descendant n$, then
from \cref{ax_distantrel_breaks} we know that if they are not directly related,
then $\cxynv\cc\czwmv$ breaks all filesystems, and they cannot be independent.
From the construction and destruction pairs and 
\cref{ax_directchild_breaks,ax_directparent_breaks} we also see that
even if they are directly related, either
$\cxynv\cc\czwmv$ or $\czwmv\cc\cxynv$ 
breaks all filesystems, so they again cannot be independent.
\end{proof}

\begin{myax}\label{combine_independent_commands}
The combination of independent commands works wherever the original commands work:
\begin{align*}
\forall\FS: &\cxynv\indep \czwmv \\
&\quad\wedge \cxynv\FS\neq\fsbroken \\
&\quad\wedge \czwmv\FS\neq\fsbroken \\
&\Rightarrow \cxynv\cc \czwmv\FS\neq\fsbroken.
\end{align*}
\end{myax}
\begin{proof}
We name this proposition an axiom for the same reason as above,
but it is easy to provide an indirect proof based on our filesystem model.
Assume that $\cxynv\cc \czwmv\FS=\fsbroken$.
We know $\cxynv\FS\neq\fsbroken$ so it must be applying 
$\czwmv$ that breaks it.
Applying a command can only result in a broken filesystem in three cases.
First, if the filesystem was already broken, which cannot be the case here.
Second, if the input type does not match the filesystem,
but we know $\FS(m)\in\setvx{W}$ and so
$(\cxynv\FS)(m)\in\setvx{W}$ as based on \cref{unrelated_is_independent}, $n\neq m$.
Third, if the new filesystem violates the tree property.
This again cannot be the case because we also know that $n\unrel m$
and the tree property only depends on the types of the parent and children of $m$,
which therefore cannot be changed by $\cxynv$.
\end{proof}

This result can be extended to sequences:

\begin{mylem}\label{combine_independent_sequences}
The combination of independent sequences works wherever the original sequences work:
\[ S\indep T \Rightarrow \worksc{S\cc T}{S,T}. \]
\end{mylem}
\begin{proof}
Assume that there is a filesystem $\FS$ so that
$S\FS\neq\fsbroken$ and $T\FS\neq\fsbroken$, but
$(S\cc T)\FS=\fsbroken$.

From \cref{unrelated_is_independent,ax_separate_commute} we know that
the commands in $S$ and $T$ pairwise commute, and so any sequence
that contains the commands from $S$ and $T$ and preserve their original partial order
is equivalent to $S\cc T$ on all filesystems.

Let the command that breaks $\FS$ in $T$ when applying $S\cc T$ be $t$
so that $T=T_0\cc t\cc T_1$.
It is still true that $(T_0 \cc t)\FS\neq\fsbroken$,
and by definition $(S\cc T_0)\FS\neq\fsbroken$,
but $(S\cc T_0\cc t)\FS=\fsbroken$.
Also, from above we know that $S\cc T_0\equiv T_0\cc S$
and so $(T_0 \cc S)\FS\neq\fsbroken$.

If we denote the first command in $S$ with $s_1$,
this means that $(T_0 \cc s_1)\FS\neq\fsbroken$,
which we can combine with $(T_0 \cc t)\FS\neq\fsbroken$, $t\indep s_1$ and
\cref{combine_independent_commands}
(using $T_0\FS$ as the reference filesystem)
to arrive at $(T_0 \cc s_1\cc t)\FS\neq\fsbroken$.

We can repeat this step for $s_2$, the next command in $S$,
and from 
$(T_0 \cc s_1\cc t)\FS\neq\fsbroken$
and
$(T_0 \cc s_1\cc s_2)\FS\neq\fsbroken$
arrive at
$(T_0 \cc s_1\cc s_2\cc t)\FS\neq\fsbroken$.
This can be repeated until $S$ is exhausted and we get
$(T_0 \cc S\cc t)\FS\neq\fsbroken$, which is a contradiction.
\end{proof}

Separately, we prove the following:

\begin{mylem}\label{worksinputmatch}
If $A$ and $B$ are minimal sequences, $\works{A,B}$,
and there are commands on node $n$ in both $A$ ($\cxynv\in A$) and $B$ ($\czwnv\in B$)
then their input types must match ($X=Z$).
\end{mylem}
\begin{proof}
This is easily shown using an indirect proof: if $X\neq Z$, then there is no filesystem that
either $\cxynv$ or $\czwnv$ would not break, 
and consequently $A$ and $B$ cannot work on the same filesystem.
\end{proof}

We continue by defining inverse commands and sequences
which allow us to move parts of sequences between the
condition and consequence parts of $\workssign$.

\begin{mydef}{Inverse commands and sequences.}
The inverse of command $\cxynv$ is $\cxynv^{-1} = \caaaa{Y}{X}{n}{\valvx}$
where $V_X$ is an arbitrary value from $\setvx{X}$.
We write $S^{-1}$ for the inverse of sequence $S$, which consists of the inverses of the commands in $S$
in reverse order.
\end{mydef}

\begin{mydef}{$\typeeq$.}
For two filesystems we write $\FS\typeeq\GS$
iff 
at all nodes the types of their values are the same, or they are both broken.
\end{mydef}

From the definitions we can clearly see that
\begin{mycor}\label{negneg_is_typeeq}
$\forall \FS: (S^{-1})^{-1}\FS\typeeq \FS$.
\end{mycor}

\begin{mylem}\label{r_invmove}
A common prefix of sequences can be moved to the other side of $\workssign$ by inverting it:
\begin{gather*}
\worksc{B\cc \seqset{A}}{\seqset{S}} \Rightarrow \worksc{\seqset{A}}{B^{-1}\cc \seqset{S}} \\
\worksc{\seqset{A}}{B\cc \seqset{S}} \Rightarrow \worksc{B^{-1}\cc \seqset{A}}{\seqset{S}}
\end{gather*}
\end{mylem}
\begin{proof}
This is based on the fact that in our model, unless they break a filesystem,
commands change one type of value into another, but never merge types.
That is,
sequences---as functions mapping filesystems to filesystems---are
essentially bijections over type-equality
with the only ``sink'' being $\fsbroken$:
\begin{align*}
\forall S,\FS,\GS: &S\FS\neq\fsbroken \Rightarrow \\
&(S\FS\ntypeeq S\GS \Leftrightarrow \FS\ntypeeq \GS).
\end{align*}

For the proposition
\[ \worksc{B\cc \seqset{A}}{\seqset{S}} \Rightarrow \worksc{\seqset{A}}{B^{-1}\cc \seqset{S}} \]
the rest of the proof is illustrated by the diagram below
where Dom($x$), the domain of $x$, represents the set of filesystems
the sequence $x$ does not break, or neither of the sequences in the set $x$ breaks.

\input{inverse-works} % TODO arrows too small

We see that the set of filesystems the sequences in $\seqset{A}$ do not break intersect with the range of $B$.
As $B$ is a bijection between its domain and range, we can use $B^{-1}$ to map this intersection back
onto the domain of $B$.
As $\worksc{B\cc \seqset{A}}{\seqset{S}}$ the domain of $\seqset{S}$ must be a subset of this
projected intersection.
If so, then we can use $B$ to map the domain of $\seqset{S}$, which yields the domain of $B^{-1}\cc \seqset{S}$.
As it is also a part of the domain of $A$, we get $\worksc{\seqset{A}}{B^{-1}\cc \seqset{S}}$.

The second part of the lemma can be proven in a similar way.
\end{proof}

\begin{mylem}\label{indep_prefix_combine}
The combination of sequences with a common head and independent tails 
continues to work under the same conditions:
\[ \worksc{A\cc B}{\seqset{S}} \wedge \worksc{A\cc C}{\seqset{S}} \wedge B\indep C \Rightarrow \worksc{A\cc B\cc C}{\seqset{S}} \]
\end{mylem}
\begin{proof}
Based on \cref{r_invmove} we know
$\worksc{B}{A^{-1}\cc \seqset{S}}$ and $\worksc{C}{A^{-1}\cc \seqset{S}}$,
and from \cref{combine_independent_sequences} we know
$\worksc{B\cc C}{B,C}$.
Combining these using \cref{workschained}
we get
$\worksc{B\cc C}{A^{-1}\cc \seqset{S}}$. 
Finally, applying the second line of \cref{r_invmove} yields
$\worksc{(A^{-1})^{-1}\cc B\cc C}{\seqset{S}}$, which proves our lemma 
as $\forall\FS: (A^{-1})^{-1}\FS\typeeq A\FS$ (\cref{negneg_is_typeeq}).
\end{proof}

>>>>

\input{macros}

@@

\section{Definitions}
\input{def-filesystem}
\input{def-commands}

\input{axioms}

@@

@@

\input{works-inverse}
